---
title: "Reproducible Research: Peer Assessment 1"
author: "Hai Yang"
date: "December 27, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Q1 Code for reading in the dataset and/or processing the data
```{r}
unzip(zipfile = "activity.zip")
data <- read.csv("activity.csv")
```
## Q2 Histogram of the total number of steps taken each day
```{r}
library(ggplot2)
total.steps <- tapply(data$steps, data$date, FUN = sum, na.rm = TRUE)
qplot(total.steps, binwidth = 1000, xlab = "total number of steps taken each day")
```

## Q3 Mean and median number of steps taken each day
```{r}
mean(total.steps, na.rm = TRUE)
median(total.steps, na.rm = TRUE)
```
## Q4 Time series plot of the average number of steps taken
```{r}
library(ggplot2)
averages <- aggregate(x = list(steps = data$steps), by = list(interval = data$interval), FUN = mean, na.rm = TRUE)

ggplot(data = averages, aes(x = interval, y = steps)) + geom_line() + xlab("5-minute interval") + 
    ylab("average number of steps taken")
```

## Q5 The 5-minute interval that, on average, contains the maximum number of steps
```{r}
averages[which.max(averages$steps), ]
```

## Q6 Code to describe and show a strategy for imputing missing data
```{r}
## Imputing missing values
missing <- is.na(data$steps)
# How many missing
table(missing)

# Replace each missing value with the mean value of its 5-minute interval
fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps)) 
        filled <- c(steps) else filled <- (averages[averages$interval == interval, "steps"])
    return(filled)
}
filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)
```

## Q7 Histogram of the total number of steps taken each day after missing values are imputed
```{r}
total.steps <- tapply(filled.data$steps, filled.data$date, FUN = sum)
qplot(total.steps, binwidth = 1000, xlab = "total number of steps taken each day")

mean(total.steps)
median(total.steps)
```

## Q8 Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")) 
        return("weekday") else if (day %in% c("Saturday", "Sunday")) 
        return("weekend") else stop("invalid date")
}
filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN = weekday.or.weekend)

averages <- aggregate(steps ~ interval + day, data = filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) + 
    xlab("5-minute interval") + ylab("Number of steps")
```

